<<<<<<< HEAD:fcast_r/fcast_r/vm_forecasting.Rmd
---
title: "Vitamix Forecasting"
author: "Carter Rudek"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

# Packages

```{r Setup, Packages and Environment, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

# Packages
library(pacman)
pacman::p_load(keyring, # passwords for snowflake
               tidyverse, # data manipulation
               odbc, DBI, # query snowflake
               lubridate, # date manipulation
               fpp3 # forecasting
)

```


# Preliminary Steps

```{r Snowflake}
# snowflake user id and password. Need to run individually and type in textbox.
key_set("MY_SF_UID")
```

```{r Snowflake}
key_set("MY_SF_PW")
```

```{r Connection, include = FALSE}
# snowflake connection
myconn <- DBI::dbConnect(odbc::odbc(), "Snowflake", uid = key_get("MY_SF_UID"), pwd = key_get("MY_SF_PW"))
```


```{r query, include = FALSE}
#df <- DBI::dbGetQuery(myconn, 'SELECT * FROM ANALYTICS.FORECASTING."sales_fcast_combined_v"')

#save.image("C:/Users/crudek/Github/practical/fcast_r/after_query_workspace.RData")
```

### Load data

```{r}
#save.image("C:/Users/crudek/Github/practical/fcast_analysis/fcast_budget_analysis/after_query_workspace.RData")
load("C:/Users/crudek/Github/practical/fcast_r/after_query_workspace.RData")
```


___________________________________________________________________________________________________________________________________________________________

# training models

```{r}
df_ts_all <- df %>%
  mutate(MONTH = yearmonth(lubridate::ymd(MONTH))) %>%
  select(DEP_ENT, MONTH, NET_SALES) %>%
  mutate(NET_SALES = round(NET_SALES)) %>%
  select(DEP_ENT, MONTH, NET_SALES) %>%
  group_by(DEP_ENT, MONTH) %>%
  summarise(NET_SALES = sum(NET_SALES)) %>%
  as_tsibble(index = MONTH, key = DEP_ENT)
```


```{r}
# training dataset for modeling
# we need to do this to determine which model is the best for each series
df_ts_all_train <- df_ts_all %>%
  filter_index(.~"2022 April") # hold out 14 months (20%) of data
```

# Fit Models

```{r}
fit_all <- df_ts_all_train %>%
  model(
    'ets' = ETS(NET_SALES),
    'arm' = ARIMA(NET_SALES)
  )
```


```{r}
# explore models
report(fit_all)
```


```{r}
# forecast using created models
# extract components and reshape data 
fcast_all <- fit_all %>%
  fabletools::forecast(h = 14) %>%
  as.data.frame() %>%
  select(-NET_SALES) %>%
  pivot_wider(names_from = '.model', values_from = '.mean')
```


_________________________________________________________________________________________________________________________________________________________

# forward looking models

```{r}
# no training, use all the data that we have to forecast for months that havent happened yet
df_ts_all_train2 <- df_ts_all %>%
  filter_index(.~"2023 June") # this is the month that we are running it in
```

# Fit Models

```{r}
fit_all2 <- df_ts_all_train2 %>%
  model(
    'ets' = ETS(NET_SALES),
    'arm' = ARIMA(NET_SALES)
  )
```


```{r}
# explore models
report(fit_all2)
```


```{r}
# forecast using created models
# extract components and reshape data 
fcast_all2 <- fit_all2 %>%
  fabletools::forecast(h = 6) %>% # we are running this in July. The models use data up until June 2023. we want to forecast the last 6 months of 2023
  as.data.frame() %>%
  select(-NET_SALES) %>%
  pivot_wider(names_from = '.model', values_from = '.mean')
```


________________________________________________________________________________________________________________________________________________________

# Combine new forecasts with original data

```{r}
combined_data <- df %>%
  left_join(union(fcast_all, fcast_all2), by = join_by(DEP_ENT, MONTH)) %>%
  mutate(ETS_ARM_AVG = (arm+ets)/2) %>%
  rename(SF_FCAST = FORECAST, ETS = ets, ARIMA = arm) %>%
  arrange(DEP_ENT, MONTH) 
```

# Write back to snowflake

```{r}
# make sure you are connected to database
table_id <- Id(database="ANALYTICS", schema="FORECASTING", table="sales_fcast_combined_t")
dbWriteTable(myconn, table_id, combined_data, append = F, overwrite = T, verbose = T)
```


































=======
---
title: "Vitamix Forecasting"
author: "Carter Rudek"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

# Packages

```{r Setup, Packages and Environment, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

# Packages
library(pacman)
pacman::p_load(keyring, # passwords for snowflake
               tidyverse, # data manipulation
               odbc, DBI, # query snowflake
               lubridate, # date manipulation
               fpp3 # forecasting
)

```


# Preliminary Steps

```{r Snowflake}
# snowflake user id and password. Need to run individually and type in textbox.
key_set("MY_SF_UID")
```

```{r Snowflake}
key_set("MY_SF_PW")
```

```{r Connection, include = FALSE}
# snowflake connection
myconn <- DBI::dbConnect(odbc::odbc(), "Snowflake", uid = key_get("MY_SF_UID"), pwd = key_get("MY_SF_PW"))
```


```{r query, include = FALSE}
#df <- DBI::dbGetQuery(myconn, 'SELECT * FROM ANALYTICS.FORECASTING."sales_fcast_combined_v"')

#save.image("C:/Users/crudek/Github/practical/fcast_r/after_query_workspace.RData")
```

### Load data

```{r}
#save.image("C:/Users/crudek/Github/practical/fcast_analysis/fcast_budget_analysis/after_query_workspace.RData")
load("C:/Users/crudek/Github/practical/fcast_r/after_query_workspace.RData")
```


___________________________________________________________________________________________________________________________________________________________

# training models

```{r}
df_ts_all <- df %>%
  mutate(MONTH = yearmonth(lubridate::ymd(MONTH))) %>%
  select(DEP_ENT, MONTH, NET_SALES) %>%
  mutate(NET_SALES = round(NET_SALES)) %>%
  select(DEP_ENT, MONTH, NET_SALES) %>%
  group_by(DEP_ENT, MONTH) %>%
  summarise(NET_SALES = sum(NET_SALES)) %>%
  as_tsibble(index = MONTH, key = DEP_ENT)
```


```{r}
# training dataset for modeling
# we need to do this to determine which model is the best for each series
df_ts_all_train <- df_ts_all %>%
  filter_index(.~"2022 April") # hold out 14 months (20%) of data
```

# Fit Models

```{r}
fit_all <- df_ts_all_train %>%
  model(
    'ets' = ETS(NET_SALES),
    'arm' = ARIMA(NET_SALES)
  )
```


```{r}
# explore models
report(fit_all)
```


```{r}
# forecast using created models
# extract components and reshape data 
fcast_all <- fit_all %>%
  fabletools::forecast(h = 14) %>%
  as.data.frame() %>%
  select(-NET_SALES) %>%
  pivot_wider(names_from = '.model', values_from = '.mean')
```


_________________________________________________________________________________________________________________________________________________________

# forward looking models

```{r}
# no training, use all the data that we have to forecast for months that havent happened yet
df_ts_all_train2 <- df_ts_all %>%
  filter_index(.~"2023 June") # this is the month that we are running it in
```

# Fit Models

```{r}
fit_all2 <- df_ts_all_train2 %>%
  model(
    'ets' = ETS(NET_SALES),
    'arm' = ARIMA(NET_SALES)
  )
```


```{r}
# explore models
report(fit_all2)
```


```{r}
# forecast using created models
# extract components and reshape data 
fcast_all2 <- fit_all2 %>%
  fabletools::forecast(h = 6) %>% # we are running this in July. The models use data up until June 2023. we want to forecast the last 6 months of 2023
  as.data.frame() %>%
  select(-NET_SALES) %>%
  pivot_wider(names_from = '.model', values_from = '.mean')
```


________________________________________________________________________________________________________________________________________________________

# Combine new forecasts with original data

```{r}
combined_data <- df %>%
  left_join(union(fcast_all, fcast_all2), by = join_by(DEP_ENT, MONTH)) %>%
  mutate(ETS_ARM_AVG = (arm+ets)/2) %>%
  rename(SF_FCAST = FORECAST, ETS = ets, ARIMA = arm) %>%
  arrange(DEP_ENT, MONTH) 
```

# Write back to snowflake

```{r}
# make sure you are connected to database
table_id <- Id(database="ANALYTICS", schema="FORECASTING", table="sales_fcast_combined_t")
DBI::dbWriteTable(myconn, table_id, combined_data, append = FALSE, overwrite = T, verbose = T)
```


































>>>>>>> cf5a438e729c208396c69da3f0da534695f62f6c:fcast_r/vm_forecasting.Rmd
